Left-Linear Rewriting in Adhesive Categories

### 2024-09-09 concur/presentazione/Figure/coffee-rewriting2.tex

  %%% REWRITING 
  % 
  %
  % 
  \begin{tikzpicture}[node distance=8mm, font=\small]
    % graph G0
    \node (G0) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
        % 
        \node at (0,1.0) [node, label=left:$1$] (1) {}
        edge [in=110, out=140, loop] node[above] {$\mathit{w}$} ()
        edge [in=220, out=250, loop] node[below] {$\mathit{b}$} ();

        \node at (0.7,1.0) [node, label=right:$2$] (2) {}
        edge [in=40, out=70, loop] node[above] {$\mathit{c}$} ();

        \draw[->] (1) to node[above, pos=0.4]{$s$} (2);        
      %
      \pgfBox
      \end{tikzpicture} 
    };
    \node[below] at (G0.south) {$G_0$};
    % graph G1
    \node [right=of G0] (G1) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm,, font=\scriptsize]
      % 
      \node at (0.6,1.0) [node, label=right:$12$] (12) {}
        edge [in=220, out=250, loop] node[below] {$\mathit{b}$} ()
        edge [in=40, out=70, loop] node[above] {$\mathit{c}$} ()
        edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();        
      %
      \pgfBox
      \end{tikzpicture}
    };
    \node[below] at (G1.south) {$G_1$};

    % graph G2
    \node [right=of G1] (G2) {        
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
        %
        \node at (0.4,0.5) {};
      \node at (0.6,1.0) [node, label=right:$12$] (12) {}
%        edge [in=220, out=250, loop] node[below] {$\mathit{b}$} ()
        edge [in=40, out=70, loop] node[above] {$\mathit{c}$} ()
        edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();        
      %
      \pgfBox
    \end{tikzpicture}
    };
    \node[below] at (G2.south) {$G_2$};

@@

    \path (G0) edge[->] node[trans, above] {$\rho_w$} (G1);
    \path (G1) edge[->] node[trans, above] {$\rho_b$} (G2);
    \path (G2) edge[->] node[trans, above] {$\rho_r$} (G3);    
  \end{tikzpicture}
\end{center}

>>>>

  %%% REWRITING 
  % 
  %
  % 
  \begin{tikzpicture}[node distance=8mm, font=\small]
    % graph G0
    \node (G0) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
        % 
        \node at (0,1.0) [node, label=left:$1$] (1) {}
        edge [in=110, out=140, loop] node[above] {$\mathit{m}$} ()
        edge [in=220, out=250, loop] node[below] {$\mathit{c}$} ();

        \node at (0.7,1.0) [node, label=right:$2$] (2) {}
        edge [in=40, out=70, loop] node[above] {$\mathit{u}$} ();

        \draw[->] (1) to node[above, pos=0.4]{$s$} (2);        
      %
      \pgfBox
      \end{tikzpicture} 
    };
    \node[below] at (G0.south) {$G_0$};
    % graph G1
    \node [right=of G0] (G1) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm,, font=\scriptsize]
      % 
      \node at (0.6,1.0) [node, label=right:$12$] (12) {}
        edge [in=220, out=250, loop] node[below] {$\mathit{c}$} ()
        edge [in=40, out=70, loop] node[above] {$\mathit{u}$} ()
        edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();        
      %
      \pgfBox
      \end{tikzpicture}
    };
    \node[below] at (G1.south) {$G_1$};

    % graph G2
    \node [right=of G1] (G2) {        
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
        %
        \node at (0.4,0.5) {};
      \node at (0.6,1.0) [node, label=right:$12$] (12) {}
%        edge [in=220, out=250, loop] node[below] {$\mathit{b}$} ()
        edge [in=40, out=70, loop] node[above] {$\mathit{u}$} ()
        edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();        
      %
      \pgfBox
    \end{tikzpicture}
    };
    \node[below] at (G2.south) {$G_2$};

@@

    \path (G0) edge[->] node[trans, above] {$\rho_m$} (G1);
    \path (G1) edge[->] node[trans, above] {$\rho_c$} (G2);
    \path (G2) edge[->] node[trans, above] {$\rho_s$} (G3);    
  \end{tikzpicture}
\end{center}

### 2024-09-09 concur/presentazione/pres_lunga.tex

	\pause 
	\medskip 
	The equivalence captures when steps are causally unrelated and thus could be
	executed in any order, and possibly in parallel.

	\pause 
	\medskip 
	Defining this equivalence often boils down to having a notion
	of independence between two consecutive steps.

@@

		A system is \emph{linear} if every rule is made by arrows in $\mathcal{M}$, \emph{left linear} if the left hand side of every rule is in $\mathcal{M}$.
	\end{definition}

@@

In left linear rewriting systems a rule can ``merge'' parts of the state.

\pause 
\begin{alertblock}{Motivation}
	Our long term aim is to provide a concurrent semantics for left linear DPO-rewriting systems (see \cite{baldan2017domains}). 
\end{alertblock}

@@

	We want to model a situation in which a coffee comes
	with a bag of sugar, a portion of milk and bag of cocoa.

@@

\begin{frame}{An example}
	If we want a (very sweet) coffee we can add first the milk and sugar, then cocoa. Finally we can stir it.
	\pause 
\begin{figure}
	\input{Figure/coffee-rewriting2.tex}
\end{figure}\pause

$\rho_s$ only needs sugar, so it can be anticipated before $\rho_c$, but in order to do so the presence of $\rho_m$ at the beginning is essential.
\end{frame}

@@

\begin{definition}
A rewriting system $(\X, R)$ is \emph{well switching} of independence pairs are unique and the Local Church-Rosser Theorem holds (sequentially independent derivation can be switched).
\end{definition}
\pause 
\begin{block}{Remark}
	It is possible to identify a larce class of categories on which every left-linear rewriting system enjoys the Local-Church Rosser Theorem holds (\cite{baldan2011adhesivity}).
\end{block}
\pause 
\begin{block}{Example}In the category $\catname{Graph}$, if the rules of a left-linear rewriting system only merge nodes and their left-hand sides contain no isolated nodes, then the system is well-switching.
\end{block}

@@

			\item 
			Suppose that $\dder{D}_1$ and $\dder{D}_2$ are independent, and that, in $\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(1,2)}
			\dder{D}_0 \cdot \dder{D}_2' \cdot \dder{D}_1'$, $\dder{D}_0$ and $\dder{D}_2'$ are independent.  If there
			is a switching sequence
			$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(0,1)}
			\dder{D}_1' \cdot \dder{D}_0' \cdot \dder{D}_2 \shift{(1,2)}
			\dder{D}_1' \cdot \dder{D}_2' \cdot \dder{D}_0''$ then in the last derivation $\dder{D}_1'$ and $\dder{D}_2'$ are independent.
		\end{enumerate}
	\end{lemma}
\end{frame}

@@

\begin{frame}{Risultati: Canonical Form}\justifying

	\begin{theorem}[No need of useless switches]
		Let $\der{D}$ and $\der{D}'$ be derivations If $\der{D} \shifteq[\sigma] \der{D}'$ then there is a switching sequence 
		%\begin{center}
		$\der{D} \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
		\shift{\nu_n} \der{D}'$
		%\end{center}
		consisting only of inversions for $\sigma$ (i.e. permutations $(i, i+1)$ such that $\sigma(i)>\sigma(i+1)$).
	\end{theorem}

\begin{theorem}[Canonical form]
	Let $\der{D}$ and $\der{D}'$ be derivations. If
	$\der{D} \shifteq[\sigma]\der{D}'$ then there is a switching sequence
	\begin{center}
		$\der{D} \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
		\shift{\nu_n} \der{D}'$
	\end{center}
	where for $i \in [1,n]$ we have $\nu_i  = \transp{k}$ with 
	$k = \max \{ j \mid \transp{j} \in \inv{\nu_{i,n}}\}$.
\end{theorem}
\end{frame}

@@

	In \cite{baldan2017domains} a category of domains, called \emph{weak prime domains} has been proposed as a possible concurrent semantic for systems allowing fusions. We plan to build on the results proved in this paper to show that to every left-linear DPO rewriting system it is possible to associate one of these domains, representing ``derivation up to switches''.
\end{frame}

>>>>

	\pause 
	\medskip 
	The equivalence captures when steps are causally unrelated and thus could be
	executed in \emph{any} order, and possibly in \emph{parallel}.

	\pause 
	\medskip 
	Defining this equivalence often boils down to having a notion
	of \emph{independence} between two consecutive steps.

@@

		A system is \emph{linear} if every rule is made by arrows in $\mathcal{M}$, \emph{left-linear} if the left hand side of every rule is in $\mathcal{M}$.
	\end{definition}

@@

In left-linear rewriting systems a rule can ``merge'' parts of the state.

\pause 
\begin{alertblock}{Motivation}
	Our long term aim is to provide a concurrent semantics for left-linear DPO-rewriting systems (see \cite{baldan2017domains}). 
\end{alertblock}

@@

	We want to model a situation in which a coffee comes
	with a bag of white sugar and a bag of brown sugar.

@@

\begin{frame}{An example}
	If we want a (very sweet) coffee we can add first the white sugar and then the brown one. Finally we can stir it.
	\pause 
\begin{figure}
	\input{Figure/coffee-rewriting2.tex}
\end{figure}\pause

$\rho_r$ only needs that some sugar is added, so it can be anticipated before $\rho_b$, but in order to do so the presence of $\rho_w$ at the beginning is essential.
\end{frame}

@@

\begin{definition}
A rewriting system $(\X, R)$ is \emph{well switching} if independence pairs are unique and the Local Church-Rosser Theorem holds (sequentially independent derivation can be switched).
\end{definition}
\pause 
\begin{block}{Remark}
	It is possible to identify a large class of categories on which every left-linear rewriting system enjoys the Local-Church Rosser Theorem holds (\cite{baldan2011adhesivity}).
\end{block}
\pause 
\begin{block}{Example}In the category $\catname{Graph}$, if the rules of a left-linear rewriting system only merge nodes and their left-hand sides contain no isolated nodes, then the system is well-switching.
\end{block}

@@

			\item 
			suppose that $\dder{D}_1$ and $\dder{D}_2$ are independent, and that, in $\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(1,2)}
			\dder{D}_0 \cdot \dder{D}_2' \cdot \dder{D}_1'$, $\dder{D}_0$ and $\dder{D}_2'$ are independent.  If there
			is a switching sequence
			$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(0,1)}
			\dder{D}_1' \cdot \dder{D}_0' \cdot \dder{D}_2 \shift{(1,2)}
			\dder{D}_1' \cdot \dder{D}_2' \cdot \dder{D}_0''$ then in the last derivation $\dder{D}_1'$ and $\dder{D}_2'$ are independent.
		\end{enumerate}
	\end{lemma}
\end{frame}

@@

\begin{frame}{Results: Canonical Form}\justifying

	\begin{theorem}[No need of useless switches]
		Let $\der{D}$ and $\der{D}'$ be derivations If $\der{D} \shifteq[\sigma] \der{D}'$ then there is a switching sequence 
		%\begin{center}
		$\der{D} \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
		\shift{\nu_n} \der{D}'$
		%\end{center}
		consisting only of inversions for $\sigma$ (i.e. permutations $(i, i+1)$ such that $\sigma(i)>\sigma(i+1)$).
	\end{theorem}
\pause 	
\begin{theorem}[Canonical form]
	Let $\der{D}$ and $\der{D}'$ be derivations. If
	$\der{D} \shifteq[\sigma]\der{D}'$ then there is a switching sequence
	\begin{center}
		$\der{D} \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
		\shift{\nu_n} \der{D}'$
	\end{center}
	where for $i \in [1,n]$ we have $\nu_i  = \transp{k}$ with 
	$k = \max \{ j \mid \transp{j} \in \inv{\nu_{i,n}}\}$.
\end{theorem}
\end{frame}

@@

	\smallskip \pause

	In \cite{baldan2017domains} a category of domains, called \emph{weak prime domains} has been proposed as a possible concurrent semantic for systems allowing fusions.

	\smallskip \pause We plan to build on the results proved in this paper to show that to every left-linear DPO rewriting system it is possible to associate one of these domains, representing ``derivation up to switches''.
\end{frame}
